Model-Based CubeSat Flight-Software Architecture using a Docs-as-Code approach

### 2022-12-03 research/manuscript-example/manuscript.tex
fix cite syntax in manuscript.tex

The m30ml was created based on a model-based systems engineering (MBSE) approach to systems engineering whereby models, as opposed to documents, serve as the authoritative source of truth for conducting systems engineering activities, such as the design, specification, analysis, verification & validation of a system \cite:{architecting_spacecraft}.

For the SeaLion Mission, a joint CubeSat mission between the Old Dominion University (ODU), Coast Guard Academy (USCGA), and the Air Force Institute of Technology (AFIT), the flight software team adopted an MBSE approach to flight software architecture, using a docs-as-code approach \cite:{docs_as_code}, whereby the same tools and methodologies for managing software are also used for configuration management of documentation. Applying both an MBSE and docs-as-code approach means that aspects of the mission architecture, such as stakeholder needs, user stories, and data structures pertaining to the CubeSat mission, are captured are within a model that is both human-readable and machine-queryable.

@@

SeaLion consists of three payloads for on-orbit validation.  ODU provided one payload while the USCGA and AFIT provided the other two payloads.  SeaLion is planned to fly as a secondary payload on a Northrop Grumman Antares Rocket from Wallops Flight Facility (WFF), currently scheduled for March, 2023 \cite:{sealion_cdr}.

@@

The CubeSat, originating from California Polytechnic State University in 1999, are a standardized form of nanosatellites.  Nanosatellites are satellites typically defined with a mass of less than 10 kg.  CubeSats, also known as Cube Satellites, are defined by standardized and modular architecture of 1-Unit (1U) cubes with dimensions of 10 cm X 10 cm X 10 cm with a maximum mass of 1.33 kg \cite:{cubesat_standard}.  These units can be stacked such as, for example, a 3U CubeSat with standard dimensions of 10 cm X 10 cm X 34.05 cm +/- 0.03 cm and a maximum mass of 4.00 kg \cite:{cubesat_standard}.

@@

CubeSats were initially conceived as educational tools for space systems engineering \cite:{heidt_new}.  Now, their roles have been expanded to not only just educational tools but for observation, technology demonstrations, and research that were previous monopolized by much larger satellites due to the aforementioned low cost of production and launch of these satellites.  As such, there has been increasing popularity for CubeSats ass seen by the number of launches in \ref{swartwout_data_graph} since year 2000 \cite:{swartwout_data}.

University groups especially are a large contributor in the overall number of launches of CubeSats yearly with university groups consistently maintaining plurality on total launches \cite:{swartwout_data}. Thus, a need for readily available and easily learnable tools for students especially exists for those who are new to both CubeSat design and systems engineering.

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=.5\textwidth]{swartwout_data_graph}
    \caption{Nanosatellite launch data provided by M. Swartwout \cite:{swartwout_data}.}
    \label{swartwout_data_graph}
\end{figure}

@@

The goal of the SeaLion CubeSat flight software architecture was to capture the data structures and expected behaviors of the flight software, such that it can be unambiguiously understood enough to be implemented, as well as provide full traceability and rationale for architectural elements with minimal configuration management overhead \cite:{sealion_mission_architecture}.

@@

Traditional approaches use documents as their authoritative source of truth for conducting system engineering activities \cite:{architecting_spacecraft}.  However, these documents often do not have a living relationship with other documents or to other corresponding elements; thus, changes to one document require manual changes to other documents \cite:{ibm_mbse}.  Models provide the following key advantages over document-based approaches \cite:{ibm_mbse}:

@@

Docs-as-code refers to a philosophy that team members should be writing documentation with the same tools as code \cite:{docs_as_code}.  These tools may include version control (e.g., Git), issue trackers, code tools, etc.

\begin{quoting}
    This means following the same workflows as development teams, and being integrated in the product team. It enables a culture where writers and developers both feel ownership of documentation, and work together to make it as good as possible \cite:{docs_as_code}. 
\end{quoting}

It was of interest to adopt a docs-as-code approach, as to yield the benefits of utilizing the same tools used to manage code, using version control tools (e.g., Git), for the configuration management of flight software architecture documentation, captured in a model-based approach \cite:{docs_as_code}.

@@

OML is a language that "enables defining systems engineering vocabularies and using them to describe systems \cite:{oml_language}." OML, inspired by Web Ontology Language 2 (OWL2) and the Semantic Web Rule Language (SWRL), is meant to be a more gentler and more disciplined method of afformentioned standard for use in systems engineering \cite:{oml_language}. OML was created in part since OWL2 does not conform easily to individual modelling rules without tooling support; OML is a tool to improve speed of modeling and the quality of models while in a more concise and human-friendly high-level external representation \cite:{oml_origin_and_rationale}. Thus, OML served as a useful basis to create the m30ml since both run on similar philosophies to simplify usage for the end user.

@@

The SeaLion Architecture uses m30ml to specify references, stakeholder needs, user stories, & data structures for facilitating traceability of design decisions within the Mission ConOps \cite:{mach30_git}.  The SeaLion repository is also structured as a Distributed Open Source Hardware (OSHW) Framework (DOF)-component for defining the contents of the Mission ConOps as a collection of nested subcomponents, component interfaces, and component functions for generating bill of materials (BOMs), assembly instructions, and/or usage documentation.

This is all done in an effort to execute the project via a documentation as code philosophy. A philosophy in which documentation is written with the same tools as code \cite:{docs_as_code}. This allows for workflows between development teams to be more closely integrated.

@@

For the aforementioned reasons, the SeaLion CubeSat flight software team downselected m30ml for specifying the SeaLion CubeSat flight software architecture \cite:{mach30_git}.

>>>>

The m30ml was created based on a model-based systems engineering (MBSE) approach to systems engineering whereby models, as opposed to documents, serve as the authoritative source of truth for conducting systems engineering activities, such as the design, specification, analysis, verification & validation of a system \cite{architecting_spacecraft}.

For the SeaLion Mission, a joint CubeSat mission between the Old Dominion University (ODU), Coast Guard Academy (USCGA), and the Air Force Institute of Technology (AFIT), the flight software team adopted an MBSE approach to flight software architecture, using a docs-as-code approach \cite{docs_as_code}, whereby the same tools and methodologies for managing software are also used for configuration management of documentation. Applying both an MBSE and docs-as-code approach means that aspects of the mission architecture, such as stakeholder needs, user stories, and data structures pertaining to the CubeSat mission, are captured are within a model that is both human-readable and machine-queryable.

@@

SeaLion consists of three payloads for on-orbit validation.  ODU provided one payload while the USCGA and AFIT provided the other two payloads.  SeaLion is planned to fly as a secondary payload on a Northrop Grumman Antares Rocket from Wallops Flight Facility (WFF), currently scheduled for March, 2023 \cite{sealion_cdr}.

@@

The CubeSat, originating from California Polytechnic State University in 1999, are a standardized form of nanosatellites.  Nanosatellites are satellites typically defined with a mass of less than 10 kg.  CubeSats, also known as Cube Satellites, are defined by standardized and modular architecture of 1-Unit (1U) cubes with dimensions of 10 cm X 10 cm X 10 cm with a maximum mass of 1.33 kg \cite{cubesat_standard}.  These units can be stacked such as, for example, a 3U CubeSat with standard dimensions of 10 cm X 10 cm X 34.05 cm +/- 0.03 cm and a maximum mass of 4.00 kg \cite{cubesat_standard}.

@@

CubeSats were initially conceived as educational tools for space systems engineering \cite{heidt_new}.  Now, their roles have been expanded to not only just educational tools but for observation, technology demonstrations, and research that were previous monopolized by much larger satellites due to the aforementioned low cost of production and launch of these satellites.  As such, there has been increasing popularity for CubeSats ass seen by the number of launches in \ref{swartwout_data_graph} since year 2000 \cite{swartwout_data}.

University groups especially are a large contributor in the overall number of launches of CubeSats yearly with university groups consistently maintaining plurality on total launches \cite{swartwout_data}. Thus, a need for readily available and easily learnable tools for students especially exists for those who are new to both CubeSat design and systems engineering.

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=.5\textwidth]{swartwout_data_graph}
    \caption{Nanosatellite launch data provided by M. Swartwout \cite{swartwout_data}.}
    \label{swartwout_data_graph}
\end{figure}

@@

The goal of the SeaLion CubeSat flight software architecture was to capture the data structures and expected behaviors of the flight software, such that it can be unambiguiously understood enough to be implemented, as well as provide full traceability and rationale for architectural elements with minimal configuration management overhead \cite{sealion_mission_architecture}.

@@

Traditional approaches use documents as their authoritative source of truth for conducting system engineering activities \cite{architecting_spacecraft}.  However, these documents often do not have a living relationship with other documents or to other corresponding elements; thus, changes to one document require manual changes to other documents \cite{ibm_mbse}.  Models provide the following key advantages over document-based approaches \cite{ibm_mbse}:

@@

Docs-as-code refers to a philosophy that team members should be writing documentation with the same tools as code \cite{docs_as_code}.  These tools may include version control (e.g., Git), issue trackers, code tools, etc.

\begin{quoting}
    This means following the same workflows as development teams, and being integrated in the product team. It enables a culture where writers and developers both feel ownership of documentation, and work together to make it as good as possible \cite{docs_as_code}. 
\end{quoting}

It was of interest to adopt a docs-as-code approach, as to yield the benefits of utilizing the same tools used to manage code, using version control tools (e.g., Git), for the configuration management of flight software architecture documentation, captured in a model-based approach \cite{docs_as_code}.

@@

OML is a language that "enables defining systems engineering vocabularies and using them to describe systems \cite{oml_language}." OML, inspired by Web Ontology Language 2 (OWL2) and the Semantic Web Rule Language (SWRL), is meant to be a more gentler and more disciplined method of afformentioned standard for use in systems engineering \cite{oml_language}. OML was created in part since OWL2 does not conform easily to individual modelling rules without tooling support; OML is a tool to improve speed of modeling and the quality of models while in a more concise and human-friendly high-level external representation \cite{oml_origin_and_rationale}. Thus, OML served as a useful basis to create the m30ml since both run on similar philosophies to simplify usage for the end user.

@@

The SeaLion Architecture uses m30ml to specify references, stakeholder needs, user stories, & data structures for facilitating traceability of design decisions within the Mission ConOps \cite{mach30_git}.  The SeaLion repository is also structured as a Distributed Open Source Hardware (OSHW) Framework (DOF)-component for defining the contents of the Mission ConOps as a collection of nested subcomponents, component interfaces, and component functions for generating bill of materials (BOMs), assembly instructions, and/or usage documentation.

This is all done in an effort to execute the project via a documentation as code philosophy. A philosophy in which documentation is written with the same tools as code \cite{docs_as_code}. This allows for workflows between development teams to be more closely integrated.

@@

For the aforementioned reasons, the SeaLion CubeSat flight software team downselected m30ml for specifying the SeaLion CubeSat flight software architecture \cite{mach30_git}.
